Fail Faster

### 2025-03-15 paper/paper.tex
intro

\usepackage{listings}
\usepackage{subcaption}

\lstset{
  mathescape=true,
  frame=none,
  xleftmargin=10pt,
  stepnumber=1,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  % language=haskell,
  keepspaces=true,
  tabsize=2,
  emphstyle={\bf},
  % commentstyle=\it\color{dkgreen},
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keywordstyle=\bfseries\ttfamily,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  % morecomment=[l]\%,
  % moredelim=**[is][\color{dkgreen}]{@}{@}
}

@@

\jwc{``Property based testing is a race against time...''}

Property-based testing (PBT) is a widely-used testing framework\hg{PBT isn't a
framework, it's a ``technique'' or ``approach'' --- Base\_QuickCheck is a framework} consisting of
two key components: a set of \textit{properties} that a system must satisfy, and
a way of producing
a large number of \textit{inputs} to that system. In contrast to traditional
unit testing, where test inputs are hand-written, PBT users create a random input
\textit{generator} that automatically produces inputs. Sometimes, this process
can itself be automated through the use of PBT libraries that synthesize
generators from type definitions; however, when inputs are required to maintain
invariants not communicated by the type, users must write a custom generator (or
filter out a broad class of invalid inputs).

>>>>

\usepackage{listings}
\usepackage{subcaption}
\usepackage{xspace}
\lstset{
  mathescape=true,
  frame=none,
  xleftmargin=10pt,
  stepnumber=1,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  % language=haskell,
  keepspaces=true,
  tabsize=2,
  emphstyle={\bf},
  % commentstyle=\it\color{dkgreen},
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keywordstyle=\bfseries\ttfamily,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  % morecomment=[l]\%,
  % moredelim=**[is][\color{dkgreen}]{@}{@}
}

@@

\jwc{``Property based testing is a race against time...''}
\\
\tr{Property-based testing (PBT) uses generators of random data to produce inputs to a
system under test, which is then checked against a set of properties the system is
 expected to satisfy.  In practice, PBT is used as a lightweight, ad hoc approach 
 to software correctness, aimed at detecting bugs within a short testing 
 window---typically 50 milliseconds to 30 seconds. Given this practical constraint, 
 it is 
 imperative that inputs be generated as fast as possible; in other words, 
 generators must be treated as performance-sensitive code.}

\tr{The locus for improvement in this area is \textit{generator libraries}: if programmers are 
 provided with the tools to write fast generators, then they do not themselves 
 need to be performance experts. However, existing libraries fail to achieve good performance 
 due to two major sources of inefficiency. The first is \textit{random sampling}, or the process
 of selecting random values. (Another sentence here?)
 The second is abstraction: generator libraries are implemented as embedded DSLS, which are flexible
 and expressive, but also introduce significant overhead that is not necessarily optimized away by 
 the compiler. These costs compound in the aggregate, raising a fundamental
 question: can generator libraries produce code that is both expressive \textit{and} efficient?
 }
 \\
\tr{
  % in strict functionallanguages like OCaml and Scala
We outline a fully general, two-stage approach to designing generator libraries, 
which provide high-level combinator functions with dramatically reduced 
overhead. First, \rand choice matters. We demonstrate that using a faster \rand
significantly impacts generator performance by directly optimizing an existing RNG, which
produces the same values as its unoptimized counterpart---but faster--thereby allowing us
to compare the RNG's impact on an existing generator library. Second, we 
describe a method of optimizing generator libraries using \textit{staged metaprogramming}, 
or ``staging''. Staged metaprogramming allows for code generation, which can mitigate 
performance challenges introduced by embedded DSLs. In particular, when a generator DSL 
introduces overhead for dynamic control flow, staging can specialize that control flow
and compile it statically. For example, when choosing between a list of options, a staged
generator can avoid allocating the list of options at runtime by expanding to an 
if-else. This approach is not tied to any particular PBT library---instead, it gives a recipe
for improving even SOTA PBT libraries.}
\tr{To demonstrate
the effectiveness of our technique, we apply our technique to a leading PBT library, \bq, built with performance engineering in mind.
Our version of this library, termed \name, is a direct drop-in replacement for the original that is completely
semantically equivalent: generators written using it produce exactly the same values as 
those written in the original version. However, in \name, the abstractions provided by the generator DSL are 
zero-cost, and the \rand is highly-optimized. The direct equivalence between \name and \bq
allows us to perform extremely fine-grained comparisons between generators---generators implemented with both versions of the
library will produce exactly the same sequence of values, so differences in testing performance are directly attributable
to \name's performance speedup.
}
\tr{
  To test \name's effectiveness, we run a bunch of experiments to show that \name runs faster, allocates less, etc., than
  \bq on a variety of different workloads. We demonstrate the generality of the technique by implementing \name in
  Scala, and benchmarking the same generators in the Scala port of the library. Finally, as a sanity check, we demonstrate
  that \name allows you to find bugs significantly faster ... .
}
\\
\tr{ Introduce waffle house... }
\tr{Harry's ``methodology'' paragraph...}
Property-based testing (PBT) is a widely-used testing framework\hg{PBT isn't a
framework, it's a ``technique'' or ``approach'' --- Base\_QuickCheck is a framework} consisting of
two key components: a set of \textit{properties} that a system must satisfy, and
a way of producing
a large number of \textit{inputs} to that system. In contrast to traditional
unit testing, where test inputs are hand-written, PBT users create a random input
\textit{generator} that automatically produces inputs. Sometimes, this process
can itself be automated through the use of PBT libraries that synthesize
generators from type definitions; however, when inputs are required to maintain
invariants not communicated by the type, users must write a custom generator (or
filter out a broad class of invalid inputs).
